\section{System Design}
\subsection{Replica}

To implement the coordinator logic within our distributed system, we integrated all coordinator functionalities directly into the \texttt{Replica} class using dedicated methods. Since a coordinator is inherently also a replica, this architectural choice allows any replica to react to specific messages and seamlessly transition into the coordinator role if it is the next node to be elected. 
To keep all nodes informed about the identity of the current leader, each replica stores the coordinator's index in a local variable, which is subsequently utilized to forward incoming write requests to the correct destination. Additionally, a boolean flag is provided to verify whether the replica itself is currently acting as the coordinator. When this flag is true, it triggers specific coordinator-only operations, such as broadcasting periodic heartbeat signals. To manage these signals effectively and maintain strict independent tracking, the heartbeat statuses for each replica are mapped and maintained within a dedicated data structure.

Beyond heartbeats, the coordinator is responsible for critical synchronization duties, including:
\begin{itemize}
    \item Sending \texttt{WriteOK} messages to the replicas upon reaching a consensus quorum.
    \item Broadcasting synchronization messages immediately after a successful election to align the state of the network.
\end{itemize}

The election protocol is triggered whenever a replica receives a \texttt{CoordinatorCrashed} message. Upon reception, the replica removes the failed leader from its list of available nodes and broadcasts an \texttt{ElectionStarted} message to initiate the protocol among the remaining peers. To prevent redundant or overlapping election cycles, any replica that receives this message sets an internal flag to true, thereby ignoring subsequent \texttt{ElectionStarted} signals. Replicas then remove the current coordinator from their active set and dynamically shift their operational state from \texttt{Normal Mode} to \texttt{Election Mode}. 
We decoupled these two operational behaviors to handle concurrent user requests safely during an election phase. Consequently, any write request received while the system is in \texttt{Election Mode} is buffered locally and its forwarding is deferred until the new coordinator has been successfully established.
To optimize network bandwidth and ensure that ideally only one election message circulates at a time, the election initiation is staggered based on replica identifiers. The replica with the lowest active ID is expected to start the process first. Replicas with higher IDs defer their transmission by a timeout calculated dynamically as:
\begin{equation}
\text{Timeout} = 100 \times \text{replica\_id} \text{ ms}
\end{equation}

If the lowest-ID replica crashes or fails to initiate the election within this window, the next available node in the virtual ring will take over after its respective timeout expires, and so forth. If a node is delayed but does not crash, multiple election processes might concurrently circulate in the network; however, this does not compromise system correctness, as all replicas will eventually converge on the same coordinator.
The core of the election protocol strictly follows the project specifications, with an added optimization during the post-election phase. Once a new coordinator is elected, it reverts to \texttt{Normal Mode} and updates its view of active replicas based on which nodes contributed the latest updates during the election phase. 
The coordinator then broadcasts personalized synchronization messages tailored to each individual replica's missing updates,like if a replica missed the last two updates, the synchronization message will contain exactly those two missing entries or if a replica missed only one update, the message will contain only that specific update and so on.

When a non-coordinator replica detects that a coordinator can be elected but it is not itself the chosen candidate, it initializes a timeout proportional to its own ID. If the expected synchronization message does not arrive within this window, the replica assumes the newly elected coordinator has crashed, thereby initiating a new election cycle. By employing this ID-based timeout mechanism, we ensure that only a single replica will concurrently trigger the subsequent election.
Upon successfully receiving the synchronization message, replicas switch back to \texttt{Normal Mode}, increment the system epoch, update the current coordinator ID, apply all missing updates, and clear the pending updates queue. This flow strictly guarantees the transactional property that either all correct replicas apply the update, or none of them do.
For every election message sent or forwarded, the sender expects an acknowledgment (ACK) from the recipient. If the ACK is not received within a specific timeout, the sender assumes the target replica has crashed and forwards the election message to the subsequent node in the ring. These timeouts are managed via a queue to support concurrent election tracking—a scenario that can occur if two election processes are active simultaneously.
If an ACK failure occurs during the second phase of the election (i.e., when the election message traverses the ring for the second time), the replica removes the history entry previously inserted by that crashed node during the first phase. This ensures that the coordinator receives an up-to-date topology of active nodes. Furthermore, if the crashed node happened to be the potential new coordinator, it is removed immediately, allowing the remaining replicas to locate an alternative coordinator without restarting the entire election protocol from scratch.
Finally, crashes are simulated in strict accordance with the project specifications. The system evaluates whether a crash condition has been met—and tracks the number of remaining messages before a crash occurs—every time a message type supported by the crash configuration is processed.

