\section{Implementation}
\subsection{Replica}
% \subsection{HeartBeat}
% The coordinator when is been elected it emediatly start constructing the chedulers for each replica with a interval od 1000ms that will send each invidiually the Heartbeat message to each replica. The replicas instead will start a listen timeout for receiving the the heartbeat. if it recived on time they will cancel the last timeout and start a new one for the next neratbeat. this timeout is 100ms more then the internal of the message in otder to account of the network channel delays.
% Like said in the project specification if a replica doesnot recive the heartbeat on time that means that the coordinator has crashed and so it will becan the election sending the electionStarted message to all the other replicas and in this case the message is also sent tohimself but without using the nEtwork channel because it hisself no there is not need to insert dome network delay.
% \subsection{Election protocol}
% Upon recieinving the election started the replica will remove the current coordiantor from the replcias list and like explained in the system design part only the replica with the smalles index must start the whole election process the other will wait and if they does not recive the message on time they will send it them self becuase this means that the first has crashed or is very slow for some reason.
% Upon recieinving the election messaghe the replica will immediatlry sent the ACK to the sender and checks is the election messages conitainss its information about the last update, so the epoch end the sequance number of his. if is not present it will append his infomation to the election message to the datructure by creating a new one thus maniaineing immutability. the data structure we decided to use is a map of intager u8sed of the replica id and a LastUpdate that coniains epoch ena the seqnumber of the curerent replcia. after this it biluds another election message with the new content and it will send it to the next replcia by startign a timeout for the ack.
% if the election message coniains its information then that means that the replcia can elect a new coordinator because the messages has completed the loop through all replicas. first itcheck if is already the coordiantor in case therea re multiple election message in play and thus we dont need to run the same locgi twice. then the replica will calculate the id of the coordinatgor based on the id epoch and sequence numeber of the replcias included uin the election emssage like secified in the project specificantion if the id coordiponde to his then he is the coordinator and set the flag of the coordinator to true increase the epock and will send the personalized updates to the corrisponding replicas. After that it will begin to send the heartipt and switch to normal mode and processing all the messages like earlier.
% if is not his id then he cannot be the coordinator so it will forward the unmodified election message to the next in the ring and will sart a new timeout to listen the recieve of th esynconization message that like said in the defisn section is ID depenaent, so if the timeout expiures that measn the syncornizaion messages has not been recives so the coordinaotr has crashed and so a new election must be taken place.
% If for whatever reason the ACK does not arrive on time the replica will get the id of the crashed replica, because it has not responded on time and it checks if the ack was not recived in the phase where the coordinator cannot be elected and so it has just inserted his values and if this is the case we just warward the election to the next after the one that is crashed other wise this measn that the replica hat has crashed has previously inserted it values so we remove them and forward the mesage to the next after the crash one. like explained previsousle this avoid another enitre cyle con election protocol.
% Once the replica recives the synctonization messaghe it changes its behaviour so now it can recive all messages normally it will incremente the epoch and reset the sequencenumber will apply all the rcived updates that are once again personalized and clear the list of pending updates to setuisfy the property. after that it will begin to listen for the heartbeats like normal.
% \subsection{Crash detection}
% Once a crash message arrives if the crash is of type Now the replica crashed immediately if not we defer the crash and we save the type on a local variable.
% to implement the crash detection we use a special funcion that checks if a crash message has arrived and if so it checks if the amount of message of the current tipe are already arrived or no, if so the replica crashes immediatly if not it checks if the message arrived is of the type specified in the crash and if it is it increase the message counter, of course this funzion is called in whenever a message of type in the list of the specified list aree recived.

\subsubsection{HeartBeat}
As soon as a new coordinator is elected, it immediately initializes dedicated schedulers for each replica. These schedulers are configured with an interval of $1000\text{ ms}$ to individually transmit periodic \texttt{Heartbeat} messages to every active node. 

Conversely, the replicas initialize a countdown timer to monitor the reception of these heartbeats. If a \texttt{Heartbeat} is received within the expected timeframe, the replica cancels the current timeout and schedules a new one for the next interval. To accommodate potential network channel latencies, this timeout is configured to be $100\text{ ms}$ longer than the transmission interval:
\begin{equation}
\text{Timeout}_{\text{listen}} = \text{Interval}_{\text{heartbeat}} + 100\text{ ms}
\end{equation}

In accordance with the project specifications, if a replica does not receive a heartbeat before this timer expires, it infers that the coordinator has crashed. Consequently, it initiates the election protocol by broadcasting an \texttt{ElectionStarted} message to all other replicas. In this scenario, the replica also delivers the message to itself locally—bypassing the network channel to avoid introducing unnecessary transmission delays.

\subsubsection{Election protocol}
Upon receiving an \texttt{ElectionStarted} message, a replica removes the current coordinator from its active list. As outlined in the system design, only the replica with the smallest index is authorized to immediately initiate the actual election process. The remaining replicas defer their action by a timeout calculated dynamically as:
\begin{equation}
\text{Timeout} = 100 \times \text{replica\_id} \text{ ms}
\end{equation}

if they do not detect the start of the election within their respective timeout window, they assume the primary initiator has crashed or is experiencing severe delays, and subsequently initiate the process themselves.

When a replica receives an \texttt{Election} message, it immediately replies with an \texttt{ACK} to the sender. It then inspects the message payload to verify if its own state information (specifically, its epoch and sequence number) is already present. 
\begin{itemize}
    \item \textbf{If its information is missing:} The replica appends its data to the message structure. To preserve immutability, we copy the data into a new instance of our custom data structure—a map associating each \texttt{replica\_id} (represented as an integer) with a \texttt{LastUpdate} object (containing the respective \texttt{epoch} and \texttt{sequence\_number}). Once updated, the replica packages this data into a new \texttt{Election} message, forwards it to the next node in the virtual ring, and starts a timeout tracking the recipient's \texttt{ACK}.
    \item \textbf{If its information is already present:} This indicates that the message has completed a full loop through all active replicas, meaning a new coordinator can now be elected. 
\end{itemize}

Before proceeding with the election logic, the replica first checks if it is already acting as the coordinator. This safeguard prevents redundant executions in scenarios where multiple election messages might concurrently circulate. The replica then evaluates the state of all nodes contained within the election message to determine the new coordinator's ID, prioritizing the highest epoch, sequence number, and replica ID as defined in the specifications.

If the calculated coordinator ID matches the replica's own ID:
\begin{enumerate}
    \item It promotes itself to the coordinator role by setting its coordinator flag to \texttt{true}.
    \item It increments the system epoch.
    \item It generates and sends personalized updates to the corresponding lagging replicas.
    \item Finally, it starts broadcasting heartbeats and reverts to \texttt{Normal Mode} to resume standard message processing.
\end{enumerate}

If the calculated ID belongs to another node, the replica forwards the unmodified \texttt{Election} message to the next hop in the ring. It then starts an ID-dependent timeout to await the incoming \texttt{Synchronization} message. If this timer expires without receiving the message, the replica assumes the newly elected coordinator crashed immediately after election, thereby triggering a new election cycle.

If a expected \texttt{ACK} does not arrive within the timeout, the replica identifies the unresponsive neighbor as crashed. The system then evaluates the current phase of the protocol:
\begin{itemize}
    \item If the failure occurs during the first phase of the election (where the coordinator is not yet determined and the replica has just appended its own values), the replica simply bypasses the failed node and forwards the election message to the next available successor in the ring.
    \item If the failure occurs in the second phase, it means the crashed replica had already registered its values. To prevent an entire restart of the election cycle, the replica removes the entries associated with the crashed node from the payload and forwards the modified message to the subsequent node.
\end{itemize}

Upon receiving the final \texttt{Synchronization} message, the replicas revert to \texttt{Normal Mode}. They increment the epoch, reset the local sequence number, apply the personalized updates received, and flush the queue of pending updates to satisfy the transactional consistency guarantees of the system. Normal heartbeat monitoring is then resumed.

\subsubsection{Crash detection}
When a crash command is received, its execution depends on the specified crash type. If the crash type is classified as \texttt{Now}, the replica terminates immediately. Otherwise, the crash is deferred, and its configuration parameters are stored within local variables.

To implement this deferred crash detection, we designed a dedicated evaluation function. Every time a replica receives a message of a type listed in the crash configuration, this function is invoked to:
\begin{enumerate}
    \item Check if a pending crash has been scheduled.
    \item Increment a message counter tracking the specific message type.
    \item Compare the counter against the threshold specified in the crash command.
\end{enumerate}
Once the incoming message count matches the crash trigger threshold, the replica executes an immediate shutdown.